---
title: 'Networks and Causal Inference Short Course: Part 1'
author: "Claire Pearsall, Ashley Buchanan"
date: "2026-04-30"
output: pdf_document
---

### Using Data from the Transmission Reduction Intervention Project in Athens, Greece {.unnumbered}

1.  Step 1: Download the data and load it into R Studio

This is a simulated toy dataset made for this course. It is based on data collected from the Transmission Reduction Intervention Project in Athens Greece. Public data here: <https://www.icpsr.umich.edu/web/NAHDAP/studies/39059/versions/V2/staff>

Read more: <https://pubmed.ncbi.nlm.nih.gov/27917890/>

```{r}

# -----------------------------
# 1. Read data into R
# -----------------------------

#install packages (if needed)
#install.packages(c("readxl","dplyr","ggplot2","geepack","tidyverse"))
#load packages
library(readxl)
library(dplyr)
library(knitr)
library(ggplot2)
library(geepack)
library(tidyverse)
library(cobalt)
library(purrr)
library(scales)

#will need to test on final folder structure shared - we can post to GitHub 
nodes <- read.csv("C:\\Users\\claire.pearsall\\OneDrive - University of Rhode Island\\NCIPHER - R01 MSM HIV Networks - R01 MSM HIV Networks\\Training\\TRIP_network_simulation\\TRIPsim_nodes.csv")
#edges <- read.csv("../TRIP_network_simulation/TRIPsim_edges.csv")

head(nodes)
#head(edges)

```

### Generate descriptive statistical analyses and histograms

```{r}
# -----------------------------
# 2. Descriptive Analysis
# -----------------------------

# Basic summary statistics
summary(nodes)

make_table1 <- function(data, cat_vars, num_vars) {
  
  # Categorical variables
  cat_table <- map_dfr(cat_vars, function(var) {
    data %>%
      count(.data[[var]]) %>%
      mutate(
        variable = var,
        level = as.character(.data[[var]]),
        stat = paste0(n, " (", percent(n / sum(n), accuracy = 0.1), ")")
      ) %>%
      select(variable, level, stat)
  })
  
  # Numeric variables
  num_table <- map_dfr(num_vars, function(var) {
    data %>%
      summarise(
        mean = mean(.data[[var]], na.rm = TRUE),
        sd   = sd(.data[[var]], na.rm = TRUE)
      ) %>%
      mutate(
        variable = var,
        level = "",
        stat = sprintf("%.2f (%.2f)", mean, sd)
      ) %>%
      select(variable, level, stat)
  })
  
  # Combine
  bind_rows(cat_table, num_table)
}

table1 <- make_table1(
  data = nodes,
  cat_vars = c("posneg", "share", "education", "employment", "treatment", "outcome"),
  num_vars = c("num_nn", "num_tnn", "num_utnn")
)

table1

# Frequency tables for categorical variables
nodes %>% count(posneg)
nodes %>% count(share)
nodes %>% count(education)
nodes %>% count(employment)
nodes %>% count(treatment)
nodes %>% count(outcome)

# Means of key numeric variables
nodes %>%
  summarise(
    mean_nn = mean(num_nn, na.rm = TRUE),
    mean_tnn = mean(num_tnn, na.rm = TRUE),
    mean_utnn = mean(num_utnn, na.rm = TRUE)
  )

# Plots of network-related variables

ggplot(nodes, aes(x = factor(num_nn))) +
  geom_bar() +
  scale_x_discrete(breaks = 0:9, labels = 0:9) +
  labs(title = "Distribution of Number of Nearest Neighbors",
       x = "Number of Nearest Neighbors",
       y = "Count")

ggplot(nodes, aes(x = factor(num_tnn))) +
  geom_bar() +
  scale_x_discrete(breaks = 0:4, labels = 0:4) +
  labs(title = "Distribution of Exposed Neighbors",
       x = "Number of Exposed Neighbors",
       y = "Count")

ggplot(nodes, aes(x = factor(num_utnn))) +
  geom_bar() +
  scale_x_discrete(breaks = 0:8, labels = 0:8) +
  labs(title = "Distribution of Unexposed Neighbors",
       x = "Number of Unexposed Neighbors",
       y = "Count")



```

### Analysis of the association between community alerts and any risk behavior outcome, unadjusted and adjusted for baseline covariates.

```{r}

# -----------------------------
# 3. Assess association
# -----------------------------

#Conditional RD,RR - condition on covariates in outcome model (not marginal)

#Unadjusted models
#Risk Difference

model_rd_unadj <- glm(
  outcome ~ treatment,
  data = nodes,
  family = binomial(link = "identity")
)

#Risk Ratio
model_rr_unadj <- glm(
  outcome ~ treatment,
  data = nodes,
  family = binomial(link = "log")
)

#adjusted models
#Risk Difference
# Binomial Log model did not converge - fit with GEE log-Poisson to get a robust estimator of the variance
#rescale age (per 10 year increase)
nodes$age10 <- nodes$age/10

 # model_rr <- glm(
 #   outcome ~ treatment + age10 + posneg + share + education + employment,
 #   data = nodes,
 #   family = binomial(link = "log")
 # )

gee_rr <- geeglm(
  outcome ~ treatment + age10 + posneg + date + share + education + employment,
  data = nodes,
  id = id,
  family = poisson(link = "log"),
  corstr = "independence"
)

#Risk Ratio
# Binomial Identity model did not converge - fit with GEE Normal-identity to get a robust estimator of the variance
# model_rd <- glm(
#   outcome ~ treatment + posneg + share + education + employment,
#   data = nodes,
#   family = binomial(link = "identity")
# )

gee_rd <- geeglm(
  outcome ~ treatment + age10 + posneg + date + share + education + employment,
  data = nodes,
  id = id,
  family = gaussian(link = "identity"),
  corstr = "independence"
)

#Format output into a table
# Associational RR and RD

# Unadjusted RR
b_rr_unadj <- coef(model_rr_unadj)["treatment"]
se_rr_unadj <- sqrt(vcov(model_rr_unadj)["treatment", "treatment"])
rr_unadj <- exp(b_rr_unadj)
rr_unadj_ci <- exp(b_rr_unadj + c(-1, 1) * 1.96 * se_rr_unadj)

# Unadjusted RD
b_rd_unadj <- coef(model_rd_unadj)["treatment"]
se_rd_unadj <- sqrt(vcov(model_rd_unadj)["treatment", "treatment"])
rd_unadj <- b_rd_unadj
rd_unadj_ci <- b_rd_unadj + c(-1, 1) * 1.96 * se_rd_unadj

# Adjusted RR
b_rr_adj <- coef(gee_rr)["treatment"]
se_rr_adj <- sqrt(vcov(gee_rr)["treatment", "treatment"])
rr_adj <- exp(b_rr_adj)
rr_adj_ci <- exp(b_rr_adj + c(-1, 1) * 1.96 * se_rr_adj)

# Adjusted RD
b_rd_adj <- coef(gee_rd)["treatment"]
se_rd_adj <- sqrt(vcov(gee_rd)["treatment", "treatment"])
rd_adj <- b_rd_adj
rd_adj_ci <- b_rd_adj + c(-1, 1) * 1.96 * se_rd_adj


results <- data.frame(
  model = c("Unadjusted", "Unadjusted", "Adjusted", "Adjusted"),
  measure = c("Risk ratio", "Risk difference", "Risk ratio", "Risk difference"),
  estimate = c(rr_unadj, rd_unadj, rr_adj, rd_adj),
  lower_95_ci = c(rr_unadj_ci[1], rd_unadj_ci[1], rr_adj_ci[1], rd_adj_ci[1]),
  upper_95_ci = c(rr_unadj_ci[2], rd_unadj_ci[2], rr_adj_ci[2], rd_adj_ci[2])
)

results

#Round to 3 decimal places
results$estimate <- round(results$estimate, 3)
results$lower_95_ci <- round(results$lower_95_ci, 3)
results$upper_95_ci <- round(results$upper_95_ci, 3)

results
```

### Check the positivity assumption for random violations, making tables of the exposure to the community alerts by each covariate. Based on the exposure, there are no structural violations of positivity (exposure possible for all combinations of individual and group level covariates).

```{r}
# ----------------------------------------
# 5. Assess random violations of positivity
# ----------------------------------------
# Table Function
make_combined_table <- function(data, treatment_var, covariates, digits = 2) {
  
  treatment_var <- rlang::ensym(treatment_var)
  
  # Pivot all covariates into long format
  long_data <- data %>%
    pivot_longer(
      cols = all_of(covariates),
      names_to = "variable",
      values_to = "level"
    ) %>%
    mutate(level = as.character(level))
  
  # Count by treatment + totals
  tab <- long_data %>%
    count(variable, level, !!treatment_var) %>%
    pivot_wider(
      names_from = !!treatment_var,
      values_from = n,
      values_fill = 0
    )
  
  # Rename treatment columns (assumes 0/1)
  tab <- tab %>%
    rename(
      Untreated = `0`,
      Treated   = `1`
    )
  
  # Add totals
  tab <- tab %>%
    mutate(
      Total = Untreated + Treated
    )
  
  # percentages 
  tab <- tab %>%
    group_by(variable) %>%
    mutate(
      Untreated = paste0(Untreated, " (", round(Untreated / sum(Untreated) * 100, digits), "%)"),
      Treated   = paste0(Treated, " (", round(Treated / sum(Treated) * 100, digits), "%)"),
      Total     = paste0(Total, " (", round(Total / sum(Total) * 100, digits), "%)")
    ) %>%
    ungroup()
  
  tab %>%
    arrange(variable, level)
}

table_combined <- make_combined_table(
  data = nodes,
  treatment_var = treatment,
  covariates = c(
    "posneg", "share", "education", "employment", "date"
  )
)

table_combined

make_continuous_table <- function(data, treatment_var, cont_vars, digits = 2, show_sd = TRUE) {
  
  treatment_var <- rlang::ensym(treatment_var)
  
  map_dfr(cont_vars, function(var) {
    
    var_sym <- rlang::sym(var)
    
    # Means
    untreated_mean <- data %>%
      filter(!!treatment_var == 0) %>%
      summarise(m = mean(!!var_sym, na.rm = TRUE)) %>%
      pull(m)
    
    treated_mean <- data %>%
      filter(!!treatment_var == 1) %>%
      summarise(m = mean(!!var_sym, na.rm = TRUE)) %>%
      pull(m)
    
    overall_mean <- mean(data[[var]], na.rm = TRUE)
    
    if (show_sd) {
      untreated_sd <- data %>%
        filter(!!treatment_var == 0) %>%
        summarise(s = sd(!!var_sym, na.rm = TRUE)) %>%
        pull(s)
      
      treated_sd <- data %>%
        filter(!!treatment_var == 1) %>%
        summarise(s = sd(!!var_sym, na.rm = TRUE)) %>%
        pull(s)
      
      overall_sd <- sd(data[[var]], na.rm = TRUE)
      
      tibble(
        Variable = var,
        Untreated = sprintf("%.2f (%.2f)", untreated_mean, untreated_sd),
        Treated   = sprintf("%.2f (%.2f)", treated_mean, treated_sd),
        Total     = sprintf("%.2f (%.2f)", overall_mean, overall_sd)
      )
      
    } else {
      
      tibble(
        Variable = var,
        Untreated = round(untreated_mean, digits),
        Treated   = round(treated_mean, digits),
        Total     = round(overall_mean, digits)
      )
    }
  })
}

table_continuous <- make_continuous_table(
  data = nodes,
  treatment_var = treatment,
  cont_vars = c("avg_posneg", "avg_share", "avg_education", "avg_employment", "avg_date")
)

table_continuous

```

\### Estimation using Inverse Probability Weighted Estimator (propensity score model)

```{r}
# -----------------------------
# 5. IPW Estimator
# -----------------------------

#Fit treatment model
mylogit <- glm(treatment ~ age10 + posneg + date + share + education + employment , data = nodes, family = "binomial")

#propensity score denominator
denom.p <- predict(mylogit,type = "response")

#stabilized weights
#numerator
numer.fit=glm(treatment~1,family=binomial(), data=nodes)
numer.p=predict(numer.fit,type="response")

#denominator

#unstabilized weights
nodes$w =ifelse(nodes$treatment==0,(1-denom.p),denom.p)
#stabilized weights
nodes$sw=ifelse(nodes$treatment==0,((1-numer.p)/(1-denom.p)),(numer.p/denom.p))

#check distribution of stabilized weights
#Mean around 1 
#Deviation from 1 means possible model mispecification or positivity issue
summary(nodes$sw)

# IPW outcome model (marginal structural model) with robust variance
# Causal Risk Ratio
# Fit with stabilized weights
gee_rr_ipw <- geeglm(
  outcome ~ treatment,
  weights = sw,
  data = nodes,
  id = id,
  family = binomial(link = "log"),
  corstr = "independence"
)

#Warning in eval(family$initialize) :
  #non-integer #successes in a binomial glm!
  #Warning OK because we weighted geeglm

b_rr <- coef(gee_rr_ipw)["treatment"]
se_rr <- sqrt(vcov(gee_rr_ipw)["treatment", "treatment"])

rr <- exp(b_rr)
rr_ci <- exp(b_rr + c(-1, 1) * 1.96 * se_rr)

# Causal risk difference
# Fit with stabilized weights
gee_rd_ipw <- geeglm(
  outcome ~ treatment,
  weights = sw,
  data = nodes,
  id = id,
  family = gaussian(link = "identity"),
  corstr = "independence"
)

b_rd <- coef(gee_rd_ipw)["treatment"]
se_rd <- sqrt(vcov(gee_rd_ipw)["treatment", "treatment"])

rd <- b_rd
rd_ci <- b_rd + c(-1, 1) * 1.96 * se_rd

results_ipw <- data.frame(
  measure = c("Risk ratio", "Risk difference"),
  estimate = c(rr, rd),
  lower_95_ci = c(rr_ci[1], rd_ci[1]),
  upper_95_ci = c(rr_ci[2], rd_ci[2])
)

results_ipw$estimate <- round(results_ipw$estimate, 3)
results_ipw$lower_95_ci <- round(results_ipw$lower_95_ci, 3)
results_ipw$upper_95_ci <- round(results_ipw$upper_95_ci, 3)

results_ipw


```

### Estimation using the G-computation (outcome model)

```{r}

# -----------------------------
# 6. G-computation
# -----------------------------

# Function for G-computation
# Save as a function so we can bootstrap to get confidence interval
# Under SUTVA and IID, we sample each individual with replacement for the boot samples

gcomp_est <- function(data) {
  n <- nrow(data)

  # original data
  all <- data
  all$interv <- -1

  # counterfactual copy with A = 0
  all.untr <- all
  all.untr$interv    <- 0
  all.untr$treatment <- 0
  all.untr$outcome  <- NA

  # counterfactual copy with A = 1
  all.tr <- all
  all.tr$interv    <- 1
  all.tr$treatment <- 1
  all.tr$outcome   <- NA

  onesample <- as.data.frame(rbind(all, all.untr, all.tr))

  # outcome model
  # Use Binomial-Logit then take function of probabilities 
  glm.obj <- glm(
    outcome ~ treatment + age10 + posneg + date + share + education + employment,
    data = onesample,
    family = binomial(link = "logit")
  )
  summary(glm.obj)

  onesample$meanY <- predict(glm.obj, newdata = onesample, type = "response")

  m1 <- onesample$meanY[onesample$interv == 1]
  m0 <- onesample$meanY[onesample$interv == 0]

  risk1 <- mean(m1)
  risk0 <- mean(m0)
  rd <- risk1 - risk0
  rr <- risk1 / risk0

  c(risk1 = risk1, risk0 = risk0, rd = rd, rr = rr)
}

# Point estimates
est <- gcomp_est(nodes)
gcomp_rd <- est[3]
gcomp_rr <- est[4]

# Bootstrap CIs
set.seed(123)
B <- 1000 #Should be at least 1000 replicates for a bootstrap confidence interval

#Note: we bootstrap individuals here ignoring the network
#There are some instances that the model does not converge
boot_out <- replicate(B, {
  idx <- sample(seq_len(nrow(nodes)), replace = TRUE)
  gcomp_est(nodes[idx, , drop = FALSE])
})

boot_out <- t(boot_out)

ci <- apply(boot_out, 2, quantile, probs = c(0.025, 0.975), na.rm = TRUE)

results <- data.frame(
  measure = c("Risk under A=1", "Risk under A=0", "Risk difference", "Risk ratio"),
  estimate = est[c("risk1", "risk0", "rd", "rr")],
  lower_95_ci = ci[1, c("risk1", "risk0", "rd", "rr")],
  upper_95_ci = ci[2, c("risk1", "risk0", "rd", "rr")]
)

results$estimate <- round(results$estimate, 3)
results$lower_95_ci <- round(results$lower_95_ci, 3)
results$upper_95_ci <- round(results$upper_95_ci, 3)

results
#save confidence intervals
gcomp_rd_ci <- c(results$lower_95_ci[1],results$upper_95_ci[1])
gcomp_rr_ci <- c(results$lower_95_ci[2],results$upper_95_ci[2])


```

### Estimation using an Augmented IPW Estimator (doubly-robust). Two estimators below - augmented IPW (DR1, Lunceford and Davidian 2004) and weighted regression estimator (DR2, Robins 2007).

```{r}
##################################################
# 7. Doubly Robust Estimators
##################################################

# Outcome model for g-formula components used in DR1
nodes_obs <- nodes
nodes_obs$interv <- -1

nodes_0 <- nodes
nodes_0$interv <- 0
nodes_0$treatment <- 0
nodes_0$outcome <- NA

nodes_1 <- nodes
nodes_1$interv <- 1
nodes_1$treatment <- 1
nodes_1$outcome <- NA

stacked <- rbind(nodes_obs, nodes_0, nodes_1)

outcome_mod <- glm(
  outcome ~ treatment + age10 + posneg + date + share + education + employment,
  data = stacked,
  family = binomial(link = "logit")
)

stacked$mean_outcome <- predict(outcome_mod, newdata = stacked, type = "response")

nodes$m0 <- stacked$mean_outcome[stacked$interv == 0]
nodes$m1 <- stacked$mean_outcome[stacked$interv == 1]
nodes$denom.p <- denom.p

# DR1: augmented IPW / Lunceford-Davidian
dr1_risk1 <- mean(
  (nodes$treatment * nodes$outcome - (nodes$treatment - nodes$denom.p) * nodes$m1) / nodes$denom.p
)

dr1_risk0 <- mean(
  ((1 - nodes$treatment) * nodes$outcome + (nodes$treatment - nodes$denom.p) * nodes$m0) / (1 - nodes$denom.p)
)

dr1_rd <- dr1_risk1 - dr1_risk0
dr1_rr <- dr1_risk1 / dr1_risk0

# DR2: weighted regression / Robins
nodes$ipw <- with(nodes, treatment / denom.p + (1 - treatment) / (1 - denom.p))

dr2_rd_mod <- lm(
  outcome ~ treatment + age10 + posneg +date + share + education + employment,
  data = nodes,
  weights = ipw
)

dr2_rr_mod <- glm(
  outcome ~ treatment + age10 + posneg +date + share + education + employment,
  data = nodes,
  weights = ipw,
  family = binomial(link = "logit")
)

#Warning in eval(family$initialize) :
  #non-integer #successes in a binomial glm!
  #Warning OK because we weighted glm

# Causal risk difference

nodes_1_dr <- nodes
nodes_1_dr$treatment <- 1

nodes_0_dr <- nodes
nodes_0_dr$treatment <- 0

dr2_rd_m1 <- predict(dr2_rd_mod, newdata = nodes_1_dr, type = "response")
dr2_rd_m0 <- predict(dr2_rd_mod, newdata = nodes_0_dr, type = "response")
dr2_rd <- mean(dr2_rd_m1) - mean(dr2_rd_m0)

dr2_rr_m1 <- predict(dr2_rr_mod, newdata = nodes_1_dr, type = "response")
dr2_rr_m0 <- predict(dr2_rr_mod, newdata = nodes_0_dr, type = "response")
dr2_risk1 <- mean(dr2_rr_m1)
dr2_risk0 <- mean(dr2_rr_m0)
dr2_rr <- dr2_risk1 / dr2_risk0

dr_results <- data.frame(
  estimator = c("DR1", "DR1", "DR2", "DR2"),
  measure = c("Risk difference", "Risk ratio", "Risk difference", "Risk ratio"),
  estimate = c(dr1_rd, dr1_rr, dr2_rd, dr2_rr)
)

dr_results

#save results

dr1_rd <- dr_results$estimate[1]
dr1_rr <- dr_results$estimate[2]
dr2_rd <- dr_results$estimate[3]
dr2_rr <- dr_results$estimate[4]

```

```{r}

##################################################
# 8. Doubly Robust Estimators - Bootstrap CIs
##################################################
# Under SUTVA and IID, we sample each individual with replacement for the boot samples
#Turn into a function for DR1 and DR2 to bootstrap to get 95% CIs

estimate_effects <- function(data) {
  # -----------------------------
  # G-computation
  # -----------------------------
  dat_obs <- data
  dat_obs$interv <- -1

  dat_0 <- data
  dat_0$interv <- 0
  dat_0$treatment <- 0
  dat_0$outcome <- NA

  dat_1 <- data
  dat_1$interv <- 1
  dat_1$treatment <- 1
  dat_1$outcome <- NA

  stacked <- rbind(dat_obs, dat_0, dat_1)

  
withCallingHandlers(
  {
    outcome_mod <- glm(
    outcome ~ treatment + age10 + date + posneg + share + education + employment,
    data = stacked,
    family = binomial(link = "logit")
        )
  },
  #suppress warning for weighting in the glm
  warning = function(w) {
    if (grepl("non-integer #successes in a binomial glm!", conditionMessage(w))) {
      invokeRestart("muffleWarning")
    }
  }
)
  
  stacked$mean_outcome <- predict(outcome_mod, newdata = stacked, type = "response")

  m1 <- stacked$mean_outcome[stacked$interv == 1]
  m0 <- stacked$mean_outcome[stacked$interv == 0]

  g_risk1 <- mean(m1)
  g_risk0 <- mean(m0)
  g_rd <- g_risk1 - g_risk0
  g_rr <- g_risk1 / g_risk0

  data$m1 <- m1
  data$m0 <- m0

  # -----------------------------
  # DR1: augmented IPW
  # w is the propensity score P(treatment = 1 | L)
  # -----------------------------
  dr1_risk1 <- mean(
    (data$treatment * data$outcome - (data$treatment - data$denom.p) * data$m1) / data$denom.p
  )

  dr1_risk0 <- mean(
    ((1 - data$treatment) * data$outcome + (data$treatment - data$denom.p) * data$m0) / (1 - data$denom.p)
  )

  dr1_rd <- dr1_risk1 - dr1_risk0
  dr1_rr <- dr1_risk1 / dr1_risk0

  # -----------------------------
  # DR2: weighted regression
  # -----------------------------
  ipw <- data$treatment / data$denom.p + (1 - data$treatment) / (1 - data$denom.p)
  data$ipw <- ipw

  wreg_rd <- lm(
    outcome ~ treatment + age10 + date + posneg + share + education + employment,
    data = data,
    weights = ipw
  )

  withCallingHandlers(
  {
    wreg_rr <- glm(
    outcome ~ treatment + age10 + date + posneg + share + education + employment,
    data = data,
    weights = ipw,
    family = binomial(link = "logit")
       )
  },
  #suppress warning for weighting in the glm
  warning = function(w) {
    if (grepl("non-integer #successes in a binomial glm!", conditionMessage(w))) {
      invokeRestart("muffleWarning")
    }
  }
)
  
  dat1 <- data
  dat1$treatment <- 1

  dat0 <- data
  dat0$treatment <- 0

  rd_m1 <- predict(wreg_rd, newdata = dat1, type = "response")
  rd_m0 <- predict(wreg_rd, newdata = dat0, type = "response")
  dr2_rd <- mean(rd_m1) - mean(rd_m0)

  rr_m1 <- predict(wreg_rr, newdata = dat1, type = "response")
  rr_m0 <- predict(wreg_rr, newdata = dat0, type = "response")
  dr2_risk1 <- mean(rr_m1)
  dr2_risk0 <- mean(rr_m0)
  dr2_rr <- dr2_risk1 / dr2_risk0

  c(
    gcomp_rd = g_rd,
    gcomp_rr = g_rr,
    dr1_rd = dr1_rd,
    dr1_rr = dr1_rr,
    dr2_rd = dr2_rd,
    dr2_rr = dr2_rr
  )
}

set.seed(123)
B <- 1000

point_est <- estimate_effects(nodes)

boot_est <- replicate(B, {
  idx <- sample(seq_len(nrow(nodes)), replace = TRUE)
  estimate_effects(nodes[idx, , drop = FALSE])
})


boot_est <- t(boot_est)

ci <- apply(boot_est, 2, quantile, probs = c(0.025, 0.975), na.rm = TRUE)

results_dr <- data.frame(
  estimator = c("G-computation", "G-computation", "DR1", "DR1", "DR2", "DR2"),
  measure = c("Risk difference", "Risk ratio",
              "Risk difference", "Risk ratio",
              "Risk difference", "Risk ratio"),
  estimate = point_est[c("gcomp_rd", "gcomp_rr", "dr1_rd", "dr1_rr", "dr2_rd", "dr2_rr")],
  lower_95_ci = ci[1, c("gcomp_rd", "gcomp_rr", "dr1_rd", "dr1_rr", "dr2_rd", "dr2_rr")],
  upper_95_ci = ci[2, c("gcomp_rd", "gcomp_rr", "dr1_rd", "dr1_rr", "dr2_rd", "dr2_rr")]
)

results_dr$estimate <- round(results_dr$estimate, 3)
results_dr$lower_95_ci <- round(results_dr$lower_95_ci, 3)
results_dr$upper_95_ci <- round(results_dr$upper_95_ci, 3)

results_dr
dr1_rd_ci <- c(results_dr$lower_95_ci[3],results_dr$upper_95_ci[3])
dr1_rr_ci <- c(results_dr$lower_95_ci[4],results_dr$upper_95_ci[4])
dr2_rd_ci <- c(results_dr$lower_95_ci[5],results_dr$upper_95_ci[5])
dr2_rr_ci <- c(results_dr$lower_95_ci[6],results_dr$upper_95_ci[6])

# Suppressed: Warning in eval(family$initialize): non-integer #successes in a binomial glm!

```

```{r}
#Code to output results to Latex (optional)
z <- qnorm(0.975)

make_rr_row <- function(estimator, fit, term = "treatment") {
  b <- coef(fit)[term]
  se <- sqrt(vcov(fit)[term, term])
  est <- exp(b)
  ci <- exp(b + c(-1, 1) * z * se)

  data.frame(
    Estimator = estimator,
    Measure = "Risk ratio",
    Estimate = est,
    LCL = ci[1],
    UCL = ci[2]
  )
}

make_rd_row <- function(estimator, fit, term = "treatment") {
  b <- coef(fit)[term]
  se <- sqrt(vcov(fit)[term, term])
  ci <- b + c(-1, 1) * z * se

  data.frame(
    Estimator = estimator,
    Measure = "Risk difference",
    Estimate = b,
    LCL = ci[1],
    UCL = ci[2]
  )
}

make_manual_row <- function(estimator, measure, est, ci) {
  data.frame(
    Estimator = estimator,
    Measure = measure,
    Estimate = est,
    LCL = ci[1],
    UCL = ci[2]
  )
}

results <- bind_rows(
  make_rr_row("Unadjusted", model_rr_unadj),
  make_rd_row("Unadjusted", model_rd_unadj),

  make_rr_row("Covariate adjusted", gee_rr),
  make_rd_row("Covariate adjusted", gee_rd),

  make_rr_row("IPW", gee_rr_ipw),
  make_rd_row("IPW", gee_rd_ipw),

  make_manual_row("G-computation", "Risk ratio", gcomp_rr, gcomp_rr_ci),
  make_manual_row("G-computation", "Risk difference", gcomp_rd, gcomp_rd_ci),

  make_manual_row("DR1", "Risk ratio", dr1_rr, dr1_rr_ci),
  make_manual_row("DR1", "Risk difference", dr1_rd, dr1_rd_ci),

  make_manual_row("DR2", "Risk ratio", dr2_rr, dr2_rr_ci),
  make_manual_row("DR2", "Risk difference", dr2_rd, dr2_rd_ci)
)


table_wide <- results %>%
  mutate(
    Estimate = round(Estimate, 3),
    LCL = round(LCL, 3),
    UCL = round(UCL, 3),
    value = sprintf("%.3f (%.3f, %.3f)", Estimate, LCL, UCL)
  ) %>%
  select(Estimator, Measure, value) %>%
  pivot_wider(names_from = Measure, values_from = value) %>%
  select(Estimator, `Risk ratio`, `Risk difference`)

latex_tab <- kable(
  table_wide,
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  align = c("l", "l", "l")
)

slide_tex <- paste0(
  "\\begin{frame}{Effect Estimates}\n",
  "\\scriptsize\n",
  "\\centering\n",
  latex_tab,
  "\n\\end{frame}\n"
)

writeLines(slide_tex, "effect_estimates_slide.tex")
```

### Assess the fit of the propensity score model using overlap plots and standardized mean differences.

```{r}
##################################################
# 9. Assess the propensity score model
##################################################

# IPW from propensity score
nodes$ipw <- with(nodes, treatment / denom.p + (1 - treatment) / (1 - denom.p))

# Density plot of propensity scores by treatment group
ggplot(nodes, aes(x = denom.p, color = factor(treatment), fill = factor(treatment))) +
  geom_density(alpha = 0.25) +
  labs(
    x = "Propensity score",
    y = "Density",
    color = "Treatment",
    fill = "Treatment",
    title = "Propensity score overlap by treatment group"
  ) +
  theme_minimal()

#Standardized Mean Differences
bal <- bal.tab(
  treatment ~ age10 + date + posneg + share + education + employment,
  data = nodes,
  weights = nodes$ipw,
  un = TRUE,
  method = "weighting",
  estimand = "ATE"
)

print(bal)

smd_table <- bal$Balance
smd_table[, c("Diff.Un", "Diff.Adj")]

#Love Plot

love.plot(
  bal,
  stat = "mean.diffs",
  abs = TRUE,
  thresholds = c(m = 0.1),
  var.order = "unadjusted",
  title = "Standardized mean differences before and after IPW"
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
