---
title: 'Networks and Causal Inference Short Course: Part 3'
author: "TingFang Lee, Ashley Buchanan, URI Avenir Team"
date: "2026-04-30"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

### Using Data from the Transmission Reduction Intervention Project in Athens, Greece {.unnumbered}

\### Download the data and load it into R Studio

This is a simulated toy dataset made for this course. It is based on data collected from the Transmission Reduction Intervention Project in Athens Greece. Public data here: <https://www.icpsr.umich.edu/web/NAHDAP/studies/39059/versions/V2/staff>

Read more: <https://pubmed.ncbi.nlm.nih.gov/27917890/>

Load packages and function script...

```{r message=FALSE}
# -----------------------------
# 1. Read data and functions into R
# -----------------------------

# Set the working directory to the current file's folder
# install.packages("rstudioapi")
library(rstudioapi)
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
getwd() # verify the change

library(lme4)
library(plyr)
library(dplyr)
library(igraph)
library(numDeriv)
library(gtools)
library(doParallel)
library(foreach)
library(knitr)
library(kableExtra)
library(ggplot2)
library(forcats)
rm(list = ls())

no_cores <- detectCores() - 1  
registerDoParallel(cores=no_cores)  


#source("https://github.com/uri-ncipher/Nearest-Neighbor-estimators/blob/main/functions.R?raw=TRUE")
#source("IPW_functions.R")
#source("gcomp_dr_v2.R") #Ting-Fang removed warnings for bootstrap

# Source required R scripts directly from GitHub
source("https://raw.githubusercontent.com/uri-ncipher/Causal-Inference-and-Networks-Short-Course/refs/heads/main/Code/IPW_functions.R")
source("https://raw.githubusercontent.com/uri-ncipher/Causal-Inference-and-Networks-Short-Course/refs/heads/main/Code/gcomp_dr_v2.R")

```

### Data Preparation

Read in the synthetic nodes and edges data sets for creating the simulated network.

```{r}
# -----------------------------
# 2. Data Preparation
# -----------------------------

nodes <- read.csv("https://raw.githubusercontent.com/uri-ncipher/Causal-Inference-and-Networks-Short-Course/refs/heads/main/Data/TRIPsim_nodes.csv?token=GHSAT0AAAAAAD3744H66VXTAUX7FN33QHRU2QASL3Q")
edges <- read.csv("https://raw.githubusercontent.com/uri-ncipher/Causal-Inference-and-Networks-Short-Course/refs/heads/main/Data/TRIPsim_edges.csv?token=GHSAT0AAAAAAD3744H75UTOOZ6X5X7F25ZE2QASMGA")

net0=graph_from_data_frame(d=edges, vertices = nodes, directed=F)  # create network based on the nodes and edges data.
```

### Network visualization

```{r}

# -----------------------------
# 3. Network Visualizaiton
# -----------------------------

# plot the network generated above
plot(net0, vertex.size=1, vertex.label=NA, vertex.color="cadetblue3", vertex.frame.color="cadetblue3")
```

### Prepare the data for modeling

```{r}

# -----------------------------
# 3. Data Preparation for Modeling
# -----------------------------

# Save the number of individuals (nodes) and number of components
n=length(V(net0))       # number of individuals
m=components(net0)$no   # number of components

# generate a dataset with three columns, the first column is participant id (node id), 
# the second column is number of nearest neighbors for each participant (is also node degree), 
# the third column indicate which component the participant is in.
df=data.frame(id=1:n, na=unlist(lapply(1:n, num_neighbors, net=net0)),
                component=components(net0)$membership) 

# assign treatment, outcome, and baseline covariates to data
df$treatment=nodes$treatment
df$outcome=nodes$outcome
df$age=nodes$age
df$age10=nodes$age/10 #age per 10 years
df$agebin=ifelse(nodes$age>35,1,0) #age >35 years old
df$posneg=nodes$posneg
df$date=nodes$date
df$share=nodes$share
df$education=nodes$education
df$employment=nodes$employment
df$na_a= unlist(lapply(1:n, trt_neighbors, net=net0))
df$notna_a=df$na-df$na_a

# six covariates, including age
base_covariate=c("age10", "posneg", "date", "share", "education", "employment")

# averaging baseline covariates (average of nearest neighbors' baseline covariate values)
# Note: show code for averaging "age10" and "agebin" as examples. Average of other neighbor covariates already in dataset
#       Example code: df$avg_var1=unlist(lapply(1:n, avg_neighbors, net=net0, variable="var1")) # "var1" can be any covariate name
df$avg_agebin=unlist(lapply(1:n, avg_neighbors, net=net0, variable="agebin"))
df$avg_age10 =unlist(lapply(1:n, avg_neighbors, net=net0, variable="age10"))

df$avg_posneg = nodes$avg_posneg
df$avg_date = nodes$avg_date
df$avg_share = nodes$avg_share
df$avg_education = nodes$avg_education
df$avg_employment = nodes$avg_employment

avg_covariate=c("avg_age10","avg_posneg", "avg_date", "avg_share", "avg_education","avg_employment")
```

## Modeling

In this section, we will apply both IPW1 and IPW2 estimators for the average potential outcomes and corresponding causal effects. An individual's exposure/treatment and allocation strategy $\alpha$ jointly determine the average potential outcome (allocation strategy represents the probability that individuals in nearest neighbors set receiving the exposure/treatment in the counterfactual scenario, so it is a numeric value between 0 and 1). For each estimator, the R programs will output the **point estimation and the estimated variance of average potential outcomes** $\boldsymbol{\widehat{Y}(1, \alpha), \widehat{Y}(0, \alpha), \widehat{Y}(\alpha)}$ under three different allocation strategies $\alpha$ (i.e., 0.25, 0.50, and 0.75), respectively. Also, the R program will output the **point estimation and estimated variance of four causal effects: Direct** (DE), **Indirect** (IE), **Total** (TE), and **Overall effect** (OE).

Equations for calculating the four causal effects are:

- ${\widehat{DE} = \widehat{Y}(1, \alpha) - \widehat{Y}(0, \alpha)}$
- ${\widehat{IE} = \widehat{Y}(0, \alpha_0) - \widehat{Y}(0, \alpha_1)}$
- ${\widehat{TE} = \widehat{Y}(1, \alpha_0) - \widehat{Y}(0, \alpha_1)}$
- ${\widehat{OE} = \widehat{Y}(\alpha_0) - \widehat{Y}(\alpha_1)}$,

where $\alpha_0$, $\alpha_1$ both represents allocation strategies, and $\alpha_0 \neq  \alpha_1$ (Note: The associated [paper](https://arxiv.org/abs/2108.04865) compared the estimated average potential outcomes with $\alpha_1$ to $\alpha_0$ for the indirect, total, and overall effects. The code here compared the estimated average potential outcomes with $\alpha_0$ to $\alpha_1$ for the indirect, total, and overall effects). Users can set any values between 0 and 1 for $\alpha_0$ and $\alpha_1$. If given $\alpha$ is a vector, then the causal contrasts will produce pairwise comparisons in the sequential order of the values in $\alpha$. For example, if $\alpha = c(0.75, 0.25, 0.50)$, the contrasts for spillover effects will compare

- 

  (i) $\alpha_0 = 0.75$ versus $\alpha_1 = 0.25$;

- 

  (ii) $\alpha_0 = 0.75$ versus $\alpha_1 = 0.50$;

- 

  (iii) $\alpha_0 = 0.25$ versus $\alpha_1 = 0.50$.

### IPW1

Using IPW1 to estimate the average potential outcomes and causal effects under allocation strategies $\alpha$.

```{r}

# -----------------------------
# 4a. IPW1 Estimator
# -----------------------------

alpha=c(0.5, 0.4, 0.3, 0.2)  # set allocation strategies

base_covariate=c("agebin", "posneg", "date", "share", "education", "employment")
avg_covariate=c("avg_agebin","avg_posneg", "avg_date", "avg_share", "avg_education","avg_employment")
#Note: Age variable continuous removed due to convergence issue, used binary for now

IPW1 <- IPW_1_model(df, base_covariate, alpha)
IPW1
```

In the output above,

- The **section [[1]]** displays the point estimates (type = "point estimate" in the output) and estimated variances (type = "variance" in the output) for the average potential outcomes under three different allocation strategies (i.e., $\alpha = 0.25, 0.50, 0.75$) and each individual exposure ($a = 0, a=1$, and margin). "Margin" is the average potential outcomes for a particular allocation strategy, regardless of individual exposure status.
- The **section [[2]]** displays the point estimates (type = "Direct", "Indirect", "Total", "Overall" in the output) and estimated variances (type = "Var DE", "Var IE", "Var TE", "Var OE" in the output) of four causal effects: direct, indirect, total and overall effects, corresponding to particular allocation strategies $(\alpha_0$ and $\alpha_1)$. The estimated values are shown in "estimation" column in the output above.

### IPW2

Using IPW2 to estimate the average potential outcomes and causal effects under allocation strategies $\alpha$.

```{r}

# -----------------------------
# 4b. IPW2 Estimator
# -----------------------------
#change back to age per 10 year increase
base_covariate=c("age10", "posneg", "date", "share", "education", "employment")
avg_covariate=c("avg_age10","avg_posneg", "avg_date", "avg_share", "avg_education","avg_employment")

alpha=c(0.5, 0.4, 0.3, 0.2)  # set allocation strategies

formula_1=paste("cbind(na_a, notna_a)", "~", "treatment", "+", 
                paste(base_covariate, collapse = "+"), "+", 
                paste(avg_covariate, collapse = "+"))
formula_2=paste("treatment", "~", paste(base_covariate, collapse = "+"))
M1=glm(formula_1, family = binomial(link = "logit"), data=df)
M2=glm(formula_2, family = binomial(link = "logit"), data=df)

IPW2 <- IPW_2_model(df, M1, M2, alpha)
IPW2
```

In the output above,

- The **section [[1]]** displays the point estimates (type = "point estimate" in the output) and estimated variances (type = "variance" in the output) for the average potential outcomes under three different allocation strategies (i.e., $\alpha = 0.75, 0.50, 0.25$) and each individual exposure ($a = 0, a=1$, and margin). "Margin" is the average potential outcomes for a particular allocation strategy, regardless of individual exposure status.
- The **section [[2]]** displays the point estimates (type = "Direct", "Indirect", "Total", "Overall" in the output) and estimated variances (type = "Var DE", "Var IE", "Var TE", "Var OE" in the output) of four causal effects: direct, indirect, total and overall effects, corresponding to particular allocation strategies $(\alpha_0$ and $\alpha_1)$. The estimated values are shown in "estimation" column in the output above.

Now, we provide example code for an outcome-model based approach, g-computation. We use assumptions similar to IPW2 regarding the covariates for confounding from an individual's contacts in the network. We obtain 95% confidence intervals using a bootstrap on each individual and the summary information for their contacts' exposures and covariates.

```{r}

# ----------------------------------------------------------
# 4c. G-computation Estimator (similar assumptions to IPW2)
# ----------------------------------------------------------

#Obtain G-computation estimates with bootstrap CIs
#Uses similar assumptions for interference as IPW2
#Bootstrap unit is each individual and the information for their contacts

alpha <- c(0.5, 0.4, 0.3, 0.2)

res_gcomp <- gcomp_with_bootstrap(
  data = df,
  alpha = alpha,
  z_covariates = base_covariate,
  zn_covariates = avg_covariate,
  B = 1000, 
  seed = 2026
)

res_gcomp$results
```

Now, we provide example code for an doubly-robust approach, augmented IPW. We use assumptions similar to IPW2 regarding the covariates for confounding from an individual's contacts in the network. We obtain 95% confidence intervals using a bootstrap on each individual and the summary information for their contacts' exposures and covariates.

```{r}

# ----------------------------------------------------------
# 4d. Augmented IPW Estimator (doubly robust (DR1))
# ----------------------------------------------------------

#Obtain DR1 Estimates with Bootstrap CIs
#Uses similar assumptions for interference as IPW2
#Bootstrap unit is each individual and the information for their contacts
alpha=c(0.5, 0.4, 0.3, 0.2)  # set allocation strategies

res_dr <- dr_with_bootstrap(
  data = df,
  alpha = alpha,
  z_covariates = base_covariate,
  zn_covariates = avg_covariate,
  B = 1000, #how many bootstrap samples do we need? Usually 1000 for CIs, but can have less for demonstration, depends on precision needed SE ~ sqrt(p(1-p)/B)
  seed = 2026
)
res_dr$results 
```

Lastly, we perform assessment of the propensity score in IPW2 in the presence of interference.

```{r}

# ----------------------------------------------------------
# 5. Assessing the fit of the generalized propensity scores
# ----------------------------------------------------------

### Check covariate balance for generalized propensity score: Fit a model on each covariate, then include the individual exposure, group level exposure, and propensity scores, compare to a model with only the propensity scores


# IPW2 was fitted using "age10", so we check balance using "age10" (continuous covariate), not "agebin" (binary covariate)
base_covariate=c("age10", "posneg", "date", "share", "education", "employment")
avg_covariate=c("avg_age10","avg_posneg", "avg_date", "avg_share", "avg_education","avg_employment")

##### check propensity score balance for IPW2
mat_1 = as.matrix(cbind(1, df[, row.names(coef(summary(M1)))[-1]])) # neighbors' treatment
mat_2 = as.matrix(cbind(1, df[, row.names(coef(summary(M2)))[-1]])) # Individual treatment

group_ps <- plogis(mat_1 %*% coef(M1))
ind_ps <- plogis(mat_2 %*% coef(M2))
df$group_ps_ipw2 <- group_ps
df$ind_ps_ipw2 <- ind_ps #Ke, should this be 1-ps for those not exposed?

### Step 1: check individual-level propensity score balance
# Note: Compare two models w/ and w/o treatment: 
#       (1) ind_covariate ~ ind_treatment + ind_ps_ipw2; 
#       (2) ind_covariate ~ ind_ps_ipw2
ipw2_balance <- data.frame()
for (i in 1:length(base_covariate)) {
  f1 <- paste0(base_covariate[i], " ~ ind_ps_ipw2")
  f2 <- paste0(base_covariate[i], " ~ treatment + ind_ps_ipw2")

  if (base_covariate[i] == "age10"){
    model_reduced <- glm(f1, data = df, family = gaussian)
    model_full    <- glm(f2, data = df, family = gaussian("identity"))
  } else {
    model_reduced <- glm(f1, data = df, family = binomial("logit"))
    model_full    <- glm(f2, data = df, family = binomial("logit"))
  }
  res_temp <- lmtest::lrtest(model_reduced, model_full)
  
  ipw2_balance <- rbind(ipw2_balance,
                       c(f1, f2, round(res_temp$Chisq[2], 4), round(res_temp$`Pr(>Chisq)`[2], 4))
  )
}


### Step 2: check group-level propensity score balance
# Note: Compare two models w/ and w/o treatment:
#       (1) group_covariate ~ group_treatment + group_ps;
#       (2) group_covariate ~ group_ps.

for (i in 1:length(avg_covariate)) {
  f1 <- paste0(avg_covariate[i], " ~ group_ps_ipw2")
  f2 <- paste0(avg_covariate[i], " ~ cbind(na_a, notna_a) + group_ps_ipw2")
  
  model_reduced <- glm(f1, data = df)
  model_full    <- glm(f2, data = df)
  
  res_temp <- lmtest::lrtest(model_reduced, model_full)
  
  ipw2_balance <- rbind(ipw2_balance,
                       c(f1, f2, round(res_temp$Chisq[2], 4), round(res_temp$`Pr(>Chisq)`[2], 4))
  )
}
colnames(ipw2_balance) <- c("Reduced model", "Full model", "Chisq", "Pr(>Chisq)")
ipw2_balance


```

```{r}
#Optional: Format Latex Table for output of each estimation approach
#Create forest plot
library(dplyr)
library(forcats)
library(ggplot2)
library(knitr)
library(kableExtra)

z <- qnorm(0.975)

##################################################
# 1. Result object helper
##################################################
#library(conflicted)
#conflicts_prefer(dplyr::filter) # filter will now always use dplyr, even if MASS is loaded
#conflicts_prefer(dplyr::count)

make_result_object <- function(x) {
  if (is.list(x) && !is.data.frame(x)) {
    df_idx <- which(sapply(x, function(y) {
      is.data.frame(y) &&
        all(c("estimation", "alpha0", "alpha1", "type") %in% names(y))
    }))

    if (length(df_idx) == 0) {
      stop("No data frame with columns estimation, alpha0, alpha1, and type was found.")
    }

    x <- x[[df_idx[1]]]
  }

  if (!all(c("estimation", "alpha0", "alpha1", "type") %in% names(x))) {
    stop("Input must contain columns: estimation, alpha0, alpha1, type.")
  }

  est_rows <- x %>%
    filter(!grepl("^Var", type)) %>%
    mutate(
      effect_key = case_when(
        type == "Direct" ~ "DE",
        type == "Indirect" ~ "IE",
        type == "Total" ~ "TE",
        type == "Overall" ~ "OE",
        TRUE ~ type
      )
    )

  var_rows <- x %>%
    filter(grepl("^Var", type)) %>%
    mutate(
      effect_key = gsub("^Var\\s+", "", type),
      variance = estimation
    ) %>%
    select(alpha0, alpha1, effect_key, variance)

  results <- est_rows %>%
    left_join(
      var_rows,
      by = c("alpha0", "alpha1", "effect_key")
    ) %>%
    mutate(
      se = sqrt(variance),
      lower = estimation - z * se,
      upper = estimation + z * se
    ) %>%
    select(estimation, alpha0, alpha1, type, lower, upper)

  list(results = results)
}

##################################################
# 2. Standardize alpha display by estimator/effect
##################################################

prep_effect_results <- function(results, estimator_name) {
  out <- results %>%
    filter(!grepl("^Var", type)) %>%
    mutate(
      estimator = estimator_name,
      effect = case_when(
        type == "Direct" ~ "Direct",
        type == "Indirect" ~ "Indirect",
        type == "Total" ~ "Total",
        type == "Overall" ~ "Overall",
        TRUE ~ type
      )
    )

  # G-comp and DR store coverage columns reversed relative to display.
  if (estimator_name %in% c("G-comp", "DR")) {
    out <- out %>%
      mutate(
        raw_alpha1 = alpha0,
        raw_alpha0 = alpha1
      )
  } else {
    out <- out %>%
      mutate(
        raw_alpha1 = alpha1,
        raw_alpha0 = alpha0
      )
  }

  out %>%
    mutate(
      display_alpha1 = pmax(raw_alpha1, raw_alpha0),
      display_alpha0 = pmin(raw_alpha1, raw_alpha0),
      contrast = paste0("(", display_alpha1, ", ", display_alpha0, ")")
    ) %>%
    filter(
      case_when(
        effect == "Direct" ~ display_alpha1 == display_alpha0,
        effect %in% c("Indirect", "Total") ~ display_alpha1 > display_alpha0,
        effect == "Overall" ~ TRUE,
        TRUE ~ TRUE
      )
    )
}

##################################################
# 3. LaTeX table helper
##################################################

make_latex_effect_table <- function(results,
                                    estimator_name,
                                    caption = "Estimated causal effects",
                                    digits = 3) {
  table_df <- prep_effect_results(results, estimator_name) %>%
    mutate(
      effect = factor(effect, levels = c("Direct", "Indirect", "Total", "Overall")),
      estimate = round(estimation, digits),
      ci = ifelse(
        is.na(lower) | is.na(upper),
        "",
        paste0(
          "(", round(lower, digits), ", ",
          round(upper, digits), ")"
        )
      )
    ) %>%
    arrange(effect, desc(display_alpha1), desc(display_alpha0)) %>%
    select(
      Effect = effect,
      `$(\\alpha_1, \\alpha_0)$` = contrast,
      `Point estimate` = estimate,
      `95\\% CI` = ci
    )

  direct_n <- sum(table_df$Effect == "Direct")
  indirect_n <- sum(table_df$Effect == "Indirect")
  total_n <- sum(table_df$Effect == "Total")
  overall_n <- sum(table_df$Effect == "Overall")

  tab <- table_df %>%
    kable(
      format = "latex",
      booktabs = TRUE,
      escape = FALSE,
      caption = caption,
      align = c("l", "c", "c", "c")
    ) %>%
    kable_styling(latex_options = c("hold_position"))

  start <- 1

  if (direct_n > 0) {
    tab <- tab %>% pack_rows("Direct effects", start, start + direct_n - 1, italic = TRUE)
    start <- start + direct_n
  }

  if (indirect_n > 0) {
    tab <- tab %>% pack_rows("Indirect effects", start, start + indirect_n - 1, italic = TRUE)
    start <- start + indirect_n
  }

  if (total_n > 0) {
    tab <- tab %>% pack_rows("Total effects", start, start + total_n - 1, italic = TRUE)
    start <- start + total_n
  }

  if (overall_n > 0) {
    tab <- tab %>% pack_rows("Overall effects", start, start + overall_n - 1, italic = TRUE)
  }

  tab
}

##################################################
# 4. Create result objects
##################################################

res_ipw1 <- make_result_object(IPW1)
res_ipw2 <- make_result_object(IPW2)

res_gcomp <- if (exists("res_gcomp") && "results" %in% names(res_gcomp)) {
  res_gcomp
} else {
  make_result_object(gcomp_table)
}

res_dr <- if (exists("res_dr") && "results" %in% names(res_dr)) {
  res_dr
} else {
  make_result_object(dr_table)
}

##################################################
# 5. Create LaTeX tables
##################################################

ipw1_tab <- make_latex_effect_table(
  res_ipw1$results,
  estimator_name = "IPW1",
  caption = "IPW1 estimates of causal effects"
)

ipw2_tab <- make_latex_effect_table(
  res_ipw2$results,
  estimator_name = "IPW2",
  caption = "IPW2 estimates of causal effects"
)

gcomp_tab <- make_latex_effect_table(
  res_gcomp$results,
  estimator_name = "G-comp",
  caption = "G-computation estimates of causal effects"
)

dr_tab <- make_latex_effect_table(
  res_dr$results,
  estimator_name = "DR",
  caption = "Doubly robust estimates of causal effects"
)

cat(ipw1_tab, file = "ipw1_effects_table.tex")
cat(ipw2_tab, file = "ipw2_effects_table.tex")
cat(gcomp_tab, file = "gcomp_effects_table.tex")
cat(dr_tab, file = "dr_effects_table.tex")

##################################################
# 6. Forest plot data
##################################################

estimator_levels <- c("IPW1", "IPW2", "G-comp", "DR")

df <- bind_rows(
  prep_effect_results(res_ipw1$results, "IPW1"),
  prep_effect_results(res_ipw2$results, "IPW2"),
  prep_effect_results(res_gcomp$results, "G-comp"),
  prep_effect_results(res_dr$results, "DR")
) %>%
  mutate(
    estimator = factor(estimator, levels = estimator_levels),
    effect = factor(effect, levels = c("Direct", "Indirect", "Total", "Overall"))
  )

# Optional check: verify all estimator/effect/contrast combinations are present.
df %>%
  count(estimator, effect, contrast) %>%
  arrange(estimator, effect, contrast)

##################################################
# 7. Forest plot helper
##################################################

estimator_colors <- c(
  "IPW1" = "#009E73",
  "IPW2" = "#0072B2",
  "G-comp" = "#56B4E9",
  "DR" = "#E69F00"
)

estimator_shapes <- c(
  "IPW1" = 15,
  "IPW2" = 16,
  "G-comp" = 17,
  "DR" = 18
)

estimator_offsets <- c(
  "IPW1" = 0.27,
  "IPW2" = 0.09,
  "G-comp" = -0.09,
  "DR" = -0.27
)

make_effect_forest_plot <- function(plot_data, effect_name) {
  plot_data <- plot_data %>%
    filter(effect == effect_name)

  contrast_map <- plot_data %>%
    distinct(contrast, display_alpha1, display_alpha0) %>%
    arrange(desc(display_alpha1), desc(display_alpha0)) %>%
    mutate(contrast_id = row_number())

  plot_data <- plot_data %>%
    left_join(
      contrast_map,
      by = c("contrast", "display_alpha1", "display_alpha0")
    ) %>%
    mutate(
      y_pos = contrast_id + estimator_offsets[as.character(estimator)]
    )

  ggplot(
    plot_data,
    aes(color = estimator, shape = estimator)
  ) +
    geom_vline(
      xintercept = 0,
      linetype = "dashed",
      linewidth = 0.6,
      color = "gray30"
    ) +
    geom_segment(
      aes(x = lower, xend = upper, y = y_pos, yend = y_pos),
      linewidth = 0.8
    ) +
    geom_point(
      aes(x = estimation, y = y_pos),
      size = 2.2
    ) +
    scale_y_continuous(
      breaks = contrast_map$contrast_id,
      labels = contrast_map$contrast
    ) +
    scale_color_manual(values = estimator_colors, breaks = estimator_levels) +
    scale_shape_manual(values = estimator_shapes, breaks = estimator_levels) +
    labs(
      title = paste(effect_name, "Effects"),
      x = "Estimate (95% CI)",
      y = expression(paste("Allocation (", alpha[1], ", ", alpha[0], ")")),
      color = "Estimator",
      shape = "Estimator"
    ) +
    theme_minimal(base_size = 16) +
    theme(
      legend.position = "top",
      plot.title = element_text(face = "bold", hjust = 0),
      axis.text.y = element_text(size = 12),
      panel.grid.minor = element_blank(),
      panel.grid.major.y = element_blank()
    )
}

##################################################
# 8. Create four forest plots
##################################################

p_direct <- make_effect_forest_plot(df, "Direct")
p_indirect <- make_effect_forest_plot(df, "Indirect")
p_total <- make_effect_forest_plot(df, "Total")
p_overall <- make_effect_forest_plot(df, "Overall")

ggsave("spillover_forest_direct.pdf", p_direct, width = 10, height = 4.5)
ggsave("spillover_forest_indirect.pdf", p_indirect, width = 10, height = 5.5)
ggsave("spillover_forest_total.pdf", p_total, width = 10, height = 5.5)
ggsave("spillover_forest_overall.pdf", p_overall, width = 10, height = 5.5)


```
